---
title: "Confidence Interval"
output: html_notebook
---

**How likely is it that two numbers came from the same distribution?**

Suppose you have three numbers: a, b and c. Then, "b is within c of a" means the same as "a is within c of b". Both of those statements are equivalent to "the absolute value of the difference between a and b is less than c".

If somebody is $2\sigma$ away from you, what is a prob that you are $2\sigma$ away from them? => ~95%.

The above also means that there is a 95% prob. that $p$ is within $2\sigma_{\hat{p}}$ of $\hat{p}$.

```{r}
library(matrixStats)
```

## Estimating population standard deviation

```{r}
n = c(3,10,30,100)
std_est = numeric(length(n))
mse_error = numeric(length(n))
for (i in 1:length(n)) {
  samples = matrix(rnorm(1000*n[i], mean=175, sd=10), ncol=n[i])
  mse_error[i] = mean((sqrt(rowVars(samples)) - 10)^2)
}

plot(n,mse_error,'b', main='MSE STD estimate error vs sample size', lty='dashed', col='red')
abline(h=0)
```
So if our sample size is >=30 using estimated STD from a sample seems reasonable.

Standard deviation of a sampling distribution if we have >= 30 cases in a sample:

$$
\text{SE} = \sqrt{\frac{\text{Var}(\text{sample})}{n}}
$$

## Confidence Interval

$$
\bar{x} \pm z_{95\%}*\text{SE}
$$

### Finding 95% and 99% confidence intervals

Empirical proof that $\mu = 175$ is ~95% of the time in range $\bar{x} \pm z_{95\%}*\text{SE}$.
```{r}
n = 100
mu = 175
z_95 = 1.96
z_99 = 2.58
population = rnorm(1000, mean=mu, sd=10)
samples = matrix(sample(population, n*1000, replace=T), ncol=n)
p = rowMeans(samples)
se = sqrt(rowVars(samples)/n)
lb = p - z_95*se
up = p + z_95*se
in_range = lb <= mu & mu <= up
mean(in_range)
```
Same for 99% confidence:

```{r}
lb = p - z_99*se
up = p + z_99*se
in_range = lb <= mu & mu <= up
mean(in_range)
```