---
title: "dplyr"
output: html_notebook
---

```{r}
library(dplyr)
wine = read.csv('wine.csv', stringsAsFactors = F, encoding = 'UTF-8')
head(wine)
```
Removing columns from dataset:
```{r}
wine = wine[,-c(1,3)]
head(wine)
```
Creating a dataset by counting all observations grouped by country and then creating a new variable called count:
```{r}
wine %>% group_by(country) %>% summarize(count=n()) %>% arrange(desc(count))
```
Creating a new variable which contains the top 10 countries:

```{r}
top10_countries = wine %>% group_by(country) %>% summarize(count=n()) %>% slice_max(count, n=10) %>% select(country)
top10_countries
```
Changing the format from data frame to vector as.character referencing the country column
```{r}
top10_countries = as.character(top10_countries$country)
class(top10_countries)
```
Subsetting data selecting top ten countries and their points from wine:
```{r}
select_points=wine %>% 
  filter(country %in% top10_countries) %>%
  select(country, points) %>% 
  arrange(country)

ggplot(wine, aes(points,price)) + 
  geom_point() + 
  geom_smooth()
```

Creating top 15 percent cheapest wines with high rating using intersect function:

```{r}
top15percent=wine %>%
  arrange(desc(points)) %>%
  filter(points > quantile(points, prob = 0.85))

cheapest15percent=wine %>%
  arrange(price) %>%
  head(nrow(top15percent))

goodvalue = intersect(top15percent,cheapest15percent)

goodvalue
```